Simulation Learning in Computer Hex

### 2011-11-03 thesis/thesis.tex
Swapped out results to use the best individual instead of the combined.

\subsection{Evolution Strategies}
Evolution strategies (ES) is a collection of algorithms that operate on real-valued vectors. A general form of ES is given here, with some small modifications required for it to represent a biased simulation policy discussed in Section \ref{esmods}. More information on evolution strategies and evolutionary learning in general can be found in Eberhard and Shi's excellent textbook \emph{Computational Intelligence: Concepts to Implementations}\cite{eberhart2007computational}. The Evolution Strategies article on Scholarpedia.org also provides a good introduction to the topic\cite{Beyer:2007}.

@@

\subsection{The Hivemind System}
% TODO: Hivemind system

\section{Results and Conclusions}\label{results}
To test the effects of learning policy weights, the evolution process was ran for 41 generations. This took approximately 2 days using a single core of an Intel Xeon CPU running at 2.80GHz. Although evolution strategies is easy to parallelize by running multiple evaluations simultaneously, the current implementation is single-threaded. After training was completed, the 41st generation was used to form a single individual. This final individual was used to bias the simulation policy and compared to the uniform local bias, uniform local bias with a $\frac{5}{6}$th probability of playing locally. The control consisted of uniform random playouts. All players used a slightly modified UCT selection policy that included the variance term suggested by Gelly 2006\cite{gelly2006exploration}. A pure update policy was used, only the nodes directly in the path traversed by UCT were updated with the result of a simulation.

@@

Instead of just using the best individual from the last generation, the weight vector used is a linear combination of the weights from all individuals in generation 41. The weights are blended based on relative fitness, so the final weight vector will most reflect the top performer of generation 41, but also includes the weights of all individuals. In some respect all individuals have ``survived'' 41 generations of recombination and mutation, and so were included in the final learned vector.
% TODO: test combination against single best individual

\begin{figure}
	\begin{center}
		\begin{tabular}{c | c c c}
		 & default & uniform & uniform (tenuki) \\
		\hline
		uniform & 65\% & & \\
		uniform (tenuki) & 58.5\% & 45.5\% & \\
		learned & 81.5\% & 76.5\% & 79\% \\
		\end{tabular}
	\end{center}
	\caption{All-play-all tournament of the 4 Hivemind variants.}
	\label{fig:results}
\end{figure}

>>>>

\subsection{Evolution Strategies}\label{estrats}
Evolution strategies (ES) is a collection of algorithms that operate on real-valued vectors. A general form of ES is given here, with some small modifications required for it to represent a biased simulation policy discussed in Section \ref{esmods}. More information on evolution strategies and evolutionary learning in general can be found in Eberhard and Shi's excellent textbook \emph{Computational Intelligence: Concepts to Implementations}\cite{eberhart2007computational}. The Evolution Strategies article on Scholarpedia.org also provides a good introduction to the topic\cite{Beyer:2007}.

@@

\subsection{The Hivemind System}
Hivemind was designed for experiments in learning to play abstract strategy games on a regular board. The system consists of three modules: the evolutionary learning process, a Monte-Carlo based UCT search tree implementation, and a ``game tracker interface''. The tracker interface is currently implemented for the games of Hex, Go, and Tic-Tac-Toe. It is designed to be possible to implement many different games in the tracker, abstracting away the learning and search process. Hivemind is open-source software, available at \url{http://github.com/etherealmachine/hivemind}.

\subsubsection{Learning}
The learning module is responsible for driving the evolutionary search process. Currently, Hivemind uses Evolution Strategies (see Section \ref{estrats}), although previous versions experimented with Particle Swarm Optimization. As described earlier, the learning process consists of three steps, iterated once per generation: selection, mutation, and evaluation. In general, the learning module is responsible for finding a strategy (described in Section \ref{esmods}) that generates optimal results during the evaluation step. Evaluation is performed by selecting two individuals from the population and having them play a game against each other, each individual parameterized by their strategy. The individual that wins receives points. During the selection step, individuals are culled based on points received, and the process continues. To play a game, the learning module iteratively calls into the search module with the strategy, until the tracker module reports the game is over. The tracker module is also used to collect the game results.

\subsubsection{Search}
The search module is a simple generic implementation of Monte-Carlo Tree Search, using the UCT algorithm to traverse the tree. Section \ref{mcts} has a description of the MCTS process. The search module uses the Game Tracker to expand the tree and perform the random playouts. The entire source code for search is only about 600 lines of code. Using some extra information from the Game Tracker, search is configurable to use RAVE and AMAF heuristics so a single playout can provide information about the value of multiple nodes.

\subsubsection{Game Tracker}

\section{Results and Conclusions}\label{results}
To test the effects of learning policy weights, the evolution process was ran for 41 generations. This took approximately 2 days using a single core of an Intel Xeon CPU running at 2.80GHz. After training was completed, the best individual from the 41st generation was selected for comparison. This individual was used to bias the simulation policy and compared to the uniform local bias, uniform local bias with a $\frac{5}{6}$th probability of playing locally. The control consisted of uniform random playouts. All players used a slightly modified UCT selection policy that included the variance term suggested by Gelly 2006\cite{gelly2006exploration}. A pure update policy was used, only the nodes directly in the path traversed by UCT were updated with the result of a simulation.

@@

\begin{figure}
	\begin{center}
		\begin{tabular}{c | c c c}
		 & default & uniform & uniform (tenuki) \\
		\hline
		uniform & 63.50\% & & \\
		uniform (tenuki) & 64.00\% & 50.00\% & \\
		learned & 79.50\% & 81.50\% & 76.00\% \\
		\end{tabular}
	\end{center}
	\caption{All-play-all tournament of the 4 Hivemind variants.}
	\label{fig:results}
\end{figure}
% TODO: run tests using best instead of combined
